Bayesian analysis of the NESTA study of interventions against  verbal aggression online

### 2021-09-16 bayesianAnalysis/bayesianReport.tex
direct effect

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{dag <-}\StringTok{ }\KeywordTok{dagitty}\NormalTok{(}\StringTok{"}
\StringTok{  dag\{}
\StringTok{  CDS -> ADS -> IC  ons}
\StringTok{               U [unobserved]   }
\StringTok{               U -> CBS -> ABS  }
\StringTok{               U -> ABS        }
\StringTok{               U -> CDS -> ADS  }
\StringTok{               U -> ADS         }
\StringTok{               U -> CAS -> AAS    }
\StringTok{               U -> AAS                        }
\StringTok{               IC -> AAS        }
\StringTok{               IC -> CAS        }
\StringTok{               IT -> CAS        }
\StringTok{               IT -> AAS}
\StringTok{               CBS -> CDS -> CAS}
\StringTok{               ABS -> ADS -> AAS}
\StringTok{               \}"}\NormalTok{)}
\KeywordTok{set.seed}\NormalTok{(}\DecValTok{123}\NormalTok{)}
\KeywordTok{drawdag}\NormalTok{(dag)}
\end{Highlighting}
\end{Shaded}

\begin{center}\includegraphics[width=1\linewidth]{bayesianReport_files/figure-latex/dag1-1} \end{center}

@@

Crucially, all backdoor paths go through \textsf{ADS}, which then
becomes either a fork or a pipe, so all backdoor paths can be closed by
conditioning on \textsf{ADS}. Moreover there is only one directed
indirect path, it goes through \textsf{CAS}, so we should not condition
on it if we are to identify causal effect on attacks mediated by impact
on comments (unless we care about the direct effect of \textsf{IC} and
\textsf{IT} on \textsf{AAS}, but that's a separate question). This is in
line with the adjustment set identified algorithmically, and the same
move makes sense when we want to predict \textsf{CAS}.

@@

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{adjustmentSets}\NormalTok{(dag, }\DataTypeTok{exposure =} \KeywordTok{c}\NormalTok{(}\StringTok{"IC"}\NormalTok{, }\StringTok{"IT"}\NormalTok{), }\DataTypeTok{outcome =} \StringTok{"AAS"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## { ADS }
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{adjustmentSets}\NormalTok{(dag, }\DataTypeTok{exposure =} \KeywordTok{c}\NormalTok{(}\StringTok{"IC"}\NormalTok{, }\StringTok{"IT"}\NormalTok{), }\DataTypeTok{outcome =} \StringTok{"CAS"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## { ADS }
\end{verbatim}

@@

It's open season for other variables, and our decision to include them
in the model will be guided by information-theoretic criteria of
predictive power.

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{dag2 <-}\StringTok{ }\KeywordTok{dagitty}\NormalTok{(}\StringTok{"}
\StringTok{  dag\{}
\StringTok{                CDS -> ADS -> IC  ons}
\StringTok{                U [unobserved]   }
\StringTok{                U -> CBS -> ABS  }
\StringTok{                U -> ABS        }
\StringTok{                U -> CDS -> ADS  }
\StringTok{                U -> ADS         }
\StringTok{                U -> CAS -> AAS    }
\StringTok{                U -> AAS                        }
\StringTok{                IC -> AAS        }
\StringTok{                IC -> CAS        }
\StringTok{                IT -> CAS        }
\StringTok{                IT -> AAS}
\StringTok{                CBS -> CDS -> CAS}
\StringTok{                ABS -> ADS -> AAS}
\StringTok{                ABS -> AdiffS}
\StringTok{                AAS -> AdiffS}
\StringTok{                CBS -> CdiffS}
\StringTok{                CAS -> CdiffS}
\StringTok{                \}"}\NormalTok{)}
\KeywordTok{set.seed}\NormalTok{(}\DecValTok{123}\NormalTok{)}
\KeywordTok{drawdag}\NormalTok{(dag2)}
\KeywordTok{adjustmentSets}\NormalTok{(dag2, }\DataTypeTok{exposure =} \KeywordTok{c}\NormalTok{(}\StringTok{"IC"}\NormalTok{, }\StringTok{"IT"}\NormalTok{), }\DataTypeTok{outcome =} \StringTok{"AdiffS"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## { ADS }
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{adjustmentSets}\NormalTok{(dag2, }\DataTypeTok{exposure =} \KeywordTok{c}\NormalTok{(}\StringTok{"IC"}\NormalTok{, }\StringTok{"IT"}\NormalTok{), }\DataTypeTok{outcome =} \StringTok{"CdiffS"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## { ADS }
\end{verbatim}

\begin{center}\includegraphics[width=1\linewidth]{bayesianReport_files/figure-latex/dag2-1} \end{center}

@@

\begin{table}[H]
\centering\begingroup\fontsize{9}{11}\selectfont

\begin{tabular}{lrrrrrr}
\toprule
  & WAIC & SE & dWAIC & dSE & pWAIC & weight\\
\midrule
\cellcolor{gray!6}{Final} & \cellcolor{gray!6}{1184.641} & \cellcolor{gray!6}{89.47387} & \cellcolor{gray!6}{0.000000} & \cellcolor{gray!6}{NA} & \cellcolor{gray!6}{26.85809} & \cellcolor{gray!6}{0.8393664}\\
tooFar & 1189.257 & 90.39801 & 4.616272 & 2.884162 & 29.81499 & 0.0834720\\
\cellcolor{gray!6}{ADSITICCBS\_ITIC\_ADSIC} & \cellcolor{gray!6}{1189.414} & \cellcolor{gray!6}{87.19709} & \cellcolor{gray!6}{4.773490} & \cellcolor{gray!6}{5.762571} & \cellcolor{gray!6}{25.62296} & \cellcolor{gray!6}{0.0771616}\\
null & 1344.835 & 144.95033 & 160.193855 & 132.894118 & 17.16957 & 0.0000000\\
\cellcolor{gray!6}{IT} & \cellcolor{gray!6}{1345.879} & \cellcolor{gray!6}{144.64185} & \cellcolor{gray!6}{161.237934} & \cellcolor{gray!6}{132.478842} & \cellcolor{gray!6}{19.27433} & \cellcolor{gray!6}{0.0000000}\\
\addlinespace
ADSIT & 1348.714 & 142.29377 & 164.072877 & 130.485213 & 23.06652 & 0.0000000\\
\cellcolor{gray!6}{ADS} & \cellcolor{gray!6}{1349.706} & \cellcolor{gray!6}{145.13624} & \cellcolor{gray!6}{165.065051} & \cellcolor{gray!6}{133.323101} & \cellcolor{gray!6}{23.10634} & \cellcolor{gray!6}{0.0000000}\\
ADSIT\_ADSIT & 1351.571 & 154.69385 & 166.930496 & 140.429304 & 31.91506 & 0.0000000\\
\cellcolor{gray!6}{ADSITIC\_ADSIC} & \cellcolor{gray!6}{1352.972} & \cellcolor{gray!6}{154.55405} & \cellcolor{gray!6}{168.331385} & \cellcolor{gray!6}{140.054704} & \cellcolor{gray!6}{30.65579} & \cellcolor{gray!6}{0.0000000}\\
ADSITIC & 1353.603 & 147.38084 & 168.962571 & 134.581071 & 28.54365 & 0.0000000\\
\addlinespace
\cellcolor{gray!6}{ADSIC} & \cellcolor{gray!6}{1353.939} & \cellcolor{gray!6}{147.29717} & \cellcolor{gray!6}{169.298374} & \cellcolor{gray!6}{134.821036} & \cellcolor{gray!6}{26.91553} & \cellcolor{gray!6}{0.0000000}\\
ADSITIC\_ADSIC\_ADSIT & 1355.353 & 156.53444 & 170.712149 & 142.268677 & 33.80858 & 0.0000000\\
\cellcolor{gray!6}{ADSITIC\_ADSIT\_ITIC\_ADSIC} & \cellcolor{gray!6}{1355.442} & \cellcolor{gray!6}{155.55820} & \cellcolor{gray!6}{170.800602} & \cellcolor{gray!6}{140.950680} & \cellcolor{gray!6}{34.03861} & \cellcolor{gray!6}{0.0000000}\\
\bottomrule
\end{tabular}
\endgroup{}
\end{table}

@@

>>>>

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{dag <-}\StringTok{ }\KeywordTok{dagitty}\NormalTok{(}\StringTok{"}
\StringTok{  dag\{}
\StringTok{  CDS -> ADS -> IC  }
\StringTok{               U [unobserved]   }
\StringTok{               U -> CBS -> ABS  }
\StringTok{               U -> ABS        }
\StringTok{               U -> CDS -> ADS  }
\StringTok{               U -> ADS         }
\StringTok{               U -> CAS -> AAS    }
\StringTok{               U -> AAS                        }
\StringTok{               IC -> AAS        }
\StringTok{               IC -> CAS        }
\StringTok{               IT -> CAS        }
\StringTok{               IT -> AAS}
\StringTok{               CBS -> CDS -> CAS}
\StringTok{               ABS -> ADS -> AAS}
\StringTok{               \}"}\NormalTok{)}
\KeywordTok{coordinates}\NormalTok{(dag)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## $x
## AAS ABS ADS CAS CBS CDS  IC  IT   U 
##  NA  NA  NA  NA  NA  NA  NA  NA  NA 
## 
## $y
## AAS ABS ADS CAS CBS CDS  IC  IT   U 
##  NA  NA  NA  NA  NA  NA  NA  NA  NA
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{coordinates}\NormalTok{( dag ) <-}\StringTok{ }\KeywordTok{list}\NormalTok{( }\DataTypeTok{x=}\KeywordTok{c}\NormalTok{(}\DataTypeTok{CBS=}\DecValTok{0}\NormalTok{,}\DataTypeTok{ABS=}\DecValTok{0}\NormalTok{,}\DataTypeTok{CDS=}\DecValTok{1}\NormalTok{,}\DataTypeTok{ADS=}\DecValTok{1}\NormalTok{, }\DataTypeTok{CAS =} \DecValTok{2}\NormalTok{,}
                    \DataTypeTok{AAS =} \DecValTok{2}\NormalTok{, }\DataTypeTok{IT =} \FloatTok{1.5}\NormalTok{, }\DataTypeTok{IC =} \FloatTok{1.5}\NormalTok{, }\DataTypeTok{U =} \FloatTok{.5}\NormalTok{) ,}
\DataTypeTok{y=}\KeywordTok{c}\NormalTok{(}\DataTypeTok{CBS =}\DecValTok{0}\NormalTok{,}\DataTypeTok{ABS =} \DecValTok{1}\NormalTok{,}\DataTypeTok{CDS =} \DecValTok{0}\NormalTok{,}\DataTypeTok{ADS =} \DecValTok{1}\NormalTok{, }\DataTypeTok{CAS =} \DecValTok{0}\NormalTok{, }\DataTypeTok{AAS =} \DecValTok{1}\NormalTok{, }
    \DataTypeTok{IT =} \FloatTok{.3}\NormalTok{, }\DataTypeTok{IC =} \FloatTok{.7}\NormalTok{, }\DataTypeTok{U =}\NormalTok{.}\DecValTok{5}\NormalTok{) )}
\KeywordTok{drawdag}\NormalTok{(dag)}
\end{Highlighting}
\end{Shaded}

\begin{center}\includegraphics[width=0.8\linewidth]{bayesianReport_files/figure-latex/dag1-1} \end{center}

@@

Crucially, all backdoor paths go through \textsf{ADS}, which then
becomes either a fork or a pipe, so all backdoor paths can be closed by
conditioning on \textsf{ADS}. Moreover there is only one directed
indirect path, it goes through \textsf{CAS}, so we should not condition
on it if we are to identify causal effect on attacks mediated by impact
on comments (unless we care about the direct effect of \textsf{IC} and
\textsf{IT} on \textsf{AAS}, but that's a separate question). This is in
line with the adjustment set identified algorithmically.

@@

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{adjustmentSets}\NormalTok{(dag, }\DataTypeTok{exposure =} \KeywordTok{c}\NormalTok{(}\StringTok{"IC"}\NormalTok{, }\StringTok{"IT"}\NormalTok{), }\DataTypeTok{outcome =} \StringTok{"AAS"}\NormalTok{, }\DataTypeTok{type =} \StringTok{"all"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## { ADS }
## { ABS, ADS }
## { ADS, CBS }
## { ABS, ADS, CBS }
## { ADS, CDS }
## { ABS, ADS, CDS }
## { ADS, CBS, CDS }
## { ABS, ADS, CBS, CDS }
## { ADS, U }
## { ABS, ADS, U }
## { ADS, CBS, U }
## { ABS, ADS, CBS, U }
## { ADS, CDS, U }
## { ABS, ADS, CDS, U }
## { ADS, CBS, CDS, U }
## { ABS, ADS, CBS, CDS, U }
\end{verbatim}

@@

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{dag2 <-}\StringTok{ }\KeywordTok{dagitty}\NormalTok{(}\StringTok{"}
\StringTok{  dag\{}
\StringTok{                CDS -> ADS -> IC  }
\StringTok{                U [unobserved]   }
\StringTok{                U -> CBS -> ABS  }
\StringTok{                U -> ABS        }
\StringTok{                U -> CDS -> ADS  }
\StringTok{                U -> ADS         }
\StringTok{                U -> CAS -> AAS    }
\StringTok{                U -> AAS                        }
\StringTok{                IC -> AAS        }
\StringTok{                IC -> CAS        }
\StringTok{                IT -> CAS        }
\StringTok{                IT -> AAS}
\StringTok{                CBS -> CDS -> CAS}
\StringTok{                ABS -> ADS -> AAS}
\StringTok{                ABS -> AdiffS}
\StringTok{                AAS -> AdiffS}
\StringTok{                CBS -> CdiffS}
\StringTok{                CAS -> CdiffS}
\StringTok{                \}"}\NormalTok{)}
\KeywordTok{coordinates}\NormalTok{( dag2 ) <-}\StringTok{ }\KeywordTok{list}\NormalTok{( }\DataTypeTok{x=}\KeywordTok{c}\NormalTok{(}\DataTypeTok{CBS=}\DecValTok{0}\NormalTok{,}\DataTypeTok{ABS=}\DecValTok{0}\NormalTok{,}\DataTypeTok{CDS=}\DecValTok{1}\NormalTok{,}\DataTypeTok{ADS=}\DecValTok{1}\NormalTok{,}
    \DataTypeTok{CAS =} \DecValTok{2}\NormalTok{, }\DataTypeTok{AAS =} \DecValTok{2}\NormalTok{, }\DataTypeTok{IT =} \FloatTok{1.5}\NormalTok{, }\DataTypeTok{IC =} \FloatTok{1.5}\NormalTok{, }\DataTypeTok{U =} \FloatTok{.5}\NormalTok{, }\DataTypeTok{AdiffS=} \FloatTok{2.5}\NormalTok{, }\DataTypeTok{CdiffS =} \FloatTok{2.5}\NormalTok{) ,}
\DataTypeTok{y=}\KeywordTok{c}\NormalTok{(}\DataTypeTok{CBS =}\DecValTok{0}\NormalTok{,}\DataTypeTok{ABS =} \DecValTok{1}\NormalTok{,}\DataTypeTok{CDS =} \DecValTok{0}\NormalTok{,}\DataTypeTok{ADS =} \DecValTok{1}\NormalTok{, }\DataTypeTok{CAS =} \DecValTok{0}\NormalTok{, }\DataTypeTok{AAS =} \DecValTok{1}\NormalTok{, }\DataTypeTok{IT =} \FloatTok{.3}\NormalTok{, }\DataTypeTok{IC =} \FloatTok{.7}\NormalTok{, }\DataTypeTok{U =}\NormalTok{.}\DecValTok{5}\NormalTok{, }\DataTypeTok{CdiffS =} \FloatTok{-.2}\NormalTok{,  }\DataTypeTok{AdiffS =} \FloatTok{1.2}\NormalTok{) )}

\KeywordTok{drawdag}\NormalTok{(dag2)}
\KeywordTok{adjustmentSets}\NormalTok{(dag2, }\DataTypeTok{exposure =} \KeywordTok{c}\NormalTok{(}\StringTok{"IC"}\NormalTok{, }\StringTok{"IT"}\NormalTok{), }\DataTypeTok{outcome =} \StringTok{"AdiffS"}\NormalTok{, }\DataTypeTok{type =} \StringTok{"canonical"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## { ABS, ADS, CBS, CDS }
\end{verbatim}

\begin{center}\includegraphics[width=0.8\linewidth]{bayesianReport_files/figure-latex/dag2-1} \end{center}

Finding a maximal sensible set (canonical) of covariates suggests
inlcuding \textsf{CDS} and \textsf{ABS}. As already discussed, we do not
include \textsf{CDS} because of its strong correlation with
\textsf{CBS}. We also do not condition on \textsf{ABS}---not only
because it has a pretty strong correlation with another predictor
(\textsf{ADS}), but rather mainly because it is used to define the
output variable. In such a set-up, of course that a model including
\textsf{ABS} would have better predictive power, but since a
definitional connection is present, thinking that its inclusion in the
model tells us something about causality would be misled.

It's open season for the other variables (and interactions between
them), and our decision to include them in the model will be guided by
information-theoretic criteria of predictive power.

@@

\begin{tabular}{lrrrrrr}
\toprule
  & WAIC & SE & dWAIC & dSE & pWAIC & weight\\
\midrule
\cellcolor{gray!6}{Final} & \cellcolor{gray!6}{1184.641} & \cellcolor{gray!6}{89.47387} & \cellcolor{gray!6}{0.000000} & \cellcolor{gray!6}{NA} & \cellcolor{gray!6}{26.85809} & \cellcolor{gray!6}{0.8345169}\\
tooFar & 1189.123 & 90.35400 & 4.481669 & 2.882584 & 29.76051 & 0.0887673\\
\cellcolor{gray!6}{ADSITICCBS\_ITIC\_ADSIC} & \cellcolor{gray!6}{1189.414} & \cellcolor{gray!6}{87.19709} & \cellcolor{gray!6}{4.773490} & \cellcolor{gray!6}{5.762571} & \cellcolor{gray!6}{25.62296} & \cellcolor{gray!6}{0.0767158}\\
null & 1344.835 & 144.95033 & 160.193855 & 132.894118 & 17.16957 & 0.0000000\\
\cellcolor{gray!6}{IT} & \cellcolor{gray!6}{1345.879} & \cellcolor{gray!6}{144.64185} & \cellcolor{gray!6}{161.237934} & \cellcolor{gray!6}{132.478842} & \cellcolor{gray!6}{19.27433} & \cellcolor{gray!6}{0.0000000}\\
\addlinespace
ADSIT & 1348.714 & 142.29377 & 164.072877 & 130.485213 & 23.06652 & 0.0000000\\
\cellcolor{gray!6}{ADS} & \cellcolor{gray!6}{1349.706} & \cellcolor{gray!6}{145.13624} & \cellcolor{gray!6}{165.065051} & \cellcolor{gray!6}{133.323101} & \cellcolor{gray!6}{23.10634} & \cellcolor{gray!6}{0.0000000}\\
ADSIT\_ADSIT & 1350.941 & 153.88118 & 166.300056 & 139.579270 & 32.04695 & 0.0000000\\
\cellcolor{gray!6}{ADSITIC\_ADSIC} & \cellcolor{gray!6}{1352.972} & \cellcolor{gray!6}{154.55405} & \cellcolor{gray!6}{168.331385} & \cellcolor{gray!6}{140.054704} & \cellcolor{gray!6}{30.65579} & \cellcolor{gray!6}{0.0000000}\\
ADSITIC & 1353.603 & 147.38084 & 168.962571 & 134.581071 & 28.54365 & 0.0000000\\
\addlinespace
\cellcolor{gray!6}{ADSIC} & \cellcolor{gray!6}{1355.210} & \cellcolor{gray!6}{148.15271} & \cellcolor{gray!6}{170.568848} & \cellcolor{gray!6}{135.609996} & \cellcolor{gray!6}{27.22326} & \cellcolor{gray!6}{0.0000000}\\
ADSITIC\_ADSIC\_ADSIT & 1355.353 & 156.53444 & 170.712149 & 142.268677 & 33.80858 & 0.0000000\\
\cellcolor{gray!6}{ADSITIC\_ADSIT\_ITIC\_ADSIC} & \cellcolor{gray!6}{1355.442} & \cellcolor{gray!6}{155.55820} & \cellcolor{gray!6}{170.800602} & \cellcolor{gray!6}{140.950680} & \cellcolor{gray!6}{34.03861} & \cellcolor{gray!6}{0.0000000}\\
\bottomrule
\end{tabular}
\endgroup{}
\end{table}

@@

\CommentTok{#saveRDS(FinalHMC, file = "models/FinalHMC.rds")}

### 2021-09-16 bayesianAnalysis/bayesianReport2.Rmd
direct effect

\vspace{1mm}
\footnotesize
```{r dag1,echo=TRUE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "100%"}
dag <- dagitty("
  dag{
  CDS -> ADS -> IC  ons
               U [unobserved]   
               U -> CBS -> ABS  
               U -> ABS        
               U -> CDS -> ADS  
               U -> ADS         
               U -> CAS -> AAS    
               U -> AAS                        
               IC -> AAS        
               IC -> CAS        
               IT -> CAS        
               IT -> AAS
               CBS -> CDS -> CAS
               ABS -> ADS -> AAS
               }")
set.seed(123)
drawdag(dag)
```
\normalsize

@@

Crucially, all backdoor paths go through \textsf{ADS}, which then becomes either a fork or a pipe, so all backdoor paths can be closed by conditioning on \textsf{ADS}. Moreover there is only one directed indirect path, it goes through \textsf{CAS}, so we should not condition on it if we are to identify causal effect on attacks mediated by impact on comments (unless we care about the direct effect of \textsf{IC}  and \textsf{IT} on \textsf{AAS}, but that's a separate question). This is in line with the adjustment set identified algorithmically, and the same move makes sense when we want to predict \textsf{CAS}.

\vspace{1mm}
\footnotesize
```{r,echo=TRUE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "100%"}
adjustmentSets(dag, exposure = c("IC", "IT"), outcome = "AAS")
adjustmentSets(dag, exposure = c("IC", "IT"), outcome = "CAS")
```
\normalsize

It's open season for other variables, and our decision to include them in the model will be guided by information-theoretic criteria of predictive power.

@@

```{r dag2,echo=TRUE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "100%"}
dag2 <- dagitty("
  dag{
                CDS -> ADS -> IC  ons
                U [unobserved]   
                U -> CBS -> ABS  
                U -> ABS        
                U -> CDS -> ADS  
                U -> ADS         
                U -> CAS -> AAS    
                U -> AAS                        
                IC -> AAS        
                IC -> CAS        
                IT -> CAS        
                IT -> AAS
                CBS -> CDS -> CAS
                ABS -> ADS -> AAS
                ABS -> AdiffS
                AAS -> AdiffS
                CBS -> CdiffS
                CAS -> CdiffS
                }")
set.seed(123)
drawdag(dag2)
adjustmentSets(dag2, exposure = c("IC", "IT"), outcome = "AdiffS")
adjustmentSets(dag2, exposure = c("IC", "IT"), outcome = "CdiffS")
```

@@

>>>>

\vspace{1mm}
\footnotesize
```{r dag1,echo=TRUE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "80%"}
dag <- dagitty("
  dag{
  CDS -> ADS -> IC  
               U [unobserved]   
               U -> CBS -> ABS  
               U -> ABS        
               U -> CDS -> ADS  
               U -> ADS         
               U -> CAS -> AAS    
               U -> AAS                        
               IC -> AAS        
               IC -> CAS        
               IT -> CAS        
               IT -> AAS
               CBS -> CDS -> CAS
               ABS -> ADS -> AAS
               }")
coordinates(dag)
coordinates( dag ) <- list( x=c(CBS=0,ABS=0,CDS=1,ADS=1, CAS = 2,
                    AAS = 2, IT = 1.5, IC = 1.5, U = .5) ,
y=c(CBS =0,ABS = 1,CDS = 0,ADS = 1, CAS = 0, AAS = 1, 
    IT = .3, IC = .7, U =.5) )
drawdag(dag)
```
\normalsize

@@

Crucially, all backdoor paths go through \textsf{ADS}, which then becomes either a fork or a pipe, so all backdoor paths can be closed by conditioning on \textsf{ADS}. Moreover there is only one directed indirect path, it goes through \textsf{CAS}, so we should not condition on it if we are to identify causal effect on attacks mediated by impact on comments (unless we care about the direct effect of \textsf{IC}  and \textsf{IT} on \textsf{AAS}, but that's a separate question). This is in line with the adjustment set identified algorithmically.

\vspace{1mm}
\footnotesize
```{r,echo=TRUE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "100%"}
adjustmentSets(dag, exposure = c("IC", "IT"), outcome = "AAS", type = "all")
```
\normalsize

The situation is different when it comes to evaluating the *direct* effect of intervention. Then, we also need to block indirect causal paths from the intervention to the outcome.

\vspace{1mm}
\footnotesize
```{r,echo=TRUE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "100%"}
adjustmentSets(dag, exposure = c("IC", "IT"), outcome = "AAS", effect = "direct")
```
\normalsize

For such an evaluation we need to also condition on \textsf{CAS}, which is what we will do when we turn to the study of the direct effects of the interventions.

@@

```{r dag2,echo=TRUE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "80%"}
dag2 <- dagitty("
  dag{
                CDS -> ADS -> IC  
                U [unobserved]   
                U -> CBS -> ABS  
                U -> ABS        
                U -> CDS -> ADS  
                U -> ADS         
                U -> CAS -> AAS    
                U -> AAS                        
                IC -> AAS        
                IC -> CAS        
                IT -> CAS        
                IT -> AAS
                CBS -> CDS -> CAS
                ABS -> ADS -> AAS
                ABS -> AdiffS
                AAS -> AdiffS
                CBS -> CdiffS
                CAS -> CdiffS
                }")
coordinates( dag2 ) <- list( x=c(CBS=0,ABS=0,CDS=1,ADS=1,
    CAS = 2, AAS = 2, IT = 1.5, IC = 1.5, U = .5, AdiffS= 2.5, CdiffS = 2.5) ,
y=c(CBS =0,ABS = 1,CDS = 0,ADS = 1, CAS = 0, AAS = 1, IT = .3, IC = .7, U =.5, CdiffS = -.2,  AdiffS = 1.2) )

drawdag(dag2)
adjustmentSets(dag2, exposure = c("IC", "IT"), outcome = "AdiffS", type = "canonical")
```

Finding a maximal sensible set (canonical) of covariates suggests inlcuding \textsf{CDS} and \textsf{ABS}. As already discussed, we do not include \textsf{CDS} because of its strong correlation with \textsf{CBS}. We also do not condition on \textsf{ABS}---not only because it has a pretty strong correlation with another predictor (\textsf{ADS}), but rather mainly because it is used to define the output variable. In such a set-up, of course that a model including \textsf{ABS} would have better predictive power, but since a definitional connection is present, thinking that its inclusion in the model tells us something about causality  would be misled.

It's open season for the  other variables (and interactions between them), and our decision to include them in the model will be guided by information-theoretic criteria of predictive power.

@@

#saveRDS(FinalHMC, file = "models/FinalHMC.rds")
